---
title: "NYC 311 Calls"
author: "Natalie"
date: 2019-03-12
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(sf)

file_2010 <- "~/Desktop/DCL/C01/nyc_with_calls/nyc_with_calls_2010.rds"
file_2011 <- "~/Desktop/DCL/C01/nyc_with_calls/nyc_with_calls_2011.rds"
file_2012 <- "~/Desktop/DCL/C01/nyc_with_calls/nyc_with_calls_2012.rds"
file_2013 <- "~/Desktop/DCL/C01/nyc_with_calls/nyc_with_calls_2013.rds"
file_2014 <- "~/Desktop/DCL/C01/nyc_with_calls/nyc_with_calls_2014.rds"
file_2015 <- "~/Desktop/DCL/C01/nyc_with_calls/nyc_with_calls_2015.rds"
file_2016 <- "~/Desktop/DCL/C01/nyc_with_calls/nyc_with_calls_2016.rds"
file_2017 <- "~/Desktop/DCL/C01/nyc_with_calls/nyc_with_calls_2017.rds"

nyc_2010 <- read_rds(file_2010)
nyc_2011 <- read_rds(file_2011)
nyc_2012 <- read_rds(file_2012)
nyc_2013 <- read_rds(file_2013)
nyc_2014 <- read_rds(file_2014)
nyc_2015 <- read_rds(file_2015)
nyc_2016 <- read_rds(file_2016)
nyc_2017 <- read_rds(file_2017)
```

Now let's get plotting! I want to look intially for each year at the number of calls per tract versus other measures of the tract. Let's start with 2010:

```{r}
nyc_2010 %>% 
  drop_na(calls_per_person, total_pop) %>% 
  ggplot(mapping = aes(x = total_pop, y = calls_per_person)) +
  geom_point()
```
 
 Looking at this, we want to remove some outliers. Let's suppose that we are only going to look at tracts that have more than 2000 residents. 
 
```{r}
nyc_2010 %>% 
  filter(total_pop > 2000) %>% 
  ggplot(mapping = aes(x = total_pop, y = calls_per_person)) +
  geom_point()
```

Now let's measure versus other tract measurements. We have a few to use:

* Average rent
* Percent white
* Median household income

Average rent:

```{r}
nyc_2010 %>% 
  filter(total_pop > 2000) %>% 
  mutate(avg_rent = na_if(avg_rent, NaN)) %>% 
  drop_na(avg_rent, calls_per_person) %>% 
  ggplot(mapping = aes(x = avg_rent, y = calls_per_person)) +
  geom_point()
```

Percent white:

```{r}
nyc_2010 %>% 
  filter(total_pop > 2000) %>% 
  ggplot(mapping = aes(x = percent_white, y = calls_per_person)) +
  geom_point()
```

Median household income:

```{r}
nyc_2010 %>% 
  filter(total_pop > 2000) %>% 
  drop_na(calls_per_person, med_income) %>% 
  ggplot(mapping = aes(x = med_income, y = calls_per_person)) +
  geom_point()
```

What I'm most interested in is the change from years in terms of calls.

```{r}
reform_df <- function(data, year){
  data %>%
    mutate(
      year = year,
      rent_increase = nyc_2017$avg_rent > avg_rent,
      white_increase = nyc_2017$percent_white > percent_white,
      income_increase = nyc_2017$med_income > med_income
    )
}

nyc_2010 <- reform_df(nyc_2010, 2010)
nyc_2011 <- reform_df(nyc_2011, 2011)
nyc_2012 <- reform_df(nyc_2012, 2012)
nyc_2013 <- reform_df(nyc_2013, 2013)
nyc_2014 <- reform_df(nyc_2014, 2014)
nyc_2015 <- reform_df(nyc_2015, 2015)
nyc_2016 <- reform_df(nyc_2016, 2016)
nyc_2017 <- reform_df(nyc_2017, 2017)

full_nyc <-
  nyc_2010 %>% 
  rbind(nyc_2011, nyc_2012, nyc_2013, nyc_2014, nyc_2015, nyc_2016, nyc_2017)

nyc_3pts <-
  rbind(nyc_2010, nyc_2014, nyc_2017)
```

```{r}
full_nyc %>%   
  filter(total_pop > 2000, rent_increase == TRUE) %>% 
  mutate(avg_rent = na_if(avg_rent, NaN)) %>% 
  drop_na(avg_rent, calls_per_person) %>% 
  arrange(fips, year) %>% 
  ggplot(mapping = aes(x = avg_rent, y = calls_per_person)) +
  geom_point(aes(alpha = year), size = 2) +
  geom_path(aes(alpha = year), size = 1) +
  scale_alpha_continuous(breaks = seq(2010, 2017, 1), limits = c(2010, 2017))

full_nyc %>%   
  filter(total_pop > 2000, white_increase == TRUE) %>% 
  drop_na(percent_white, calls_per_person) %>% 
  arrange(fips, year) %>% 
  ggplot(mapping = aes(x = percent_white, y = calls_per_person)) +
  geom_point(aes(alpha = year), size = 2) +
  geom_path(aes(alpha = year), size = 1) +
  scale_alpha_continuous(breaks = seq(2010, 2017, 1), limits = c(2010, 2017))

full_nyc %>%   
  filter(total_pop > 2000, income_increase == TRUE) %>% 
  drop_na(med_income, calls_per_person) %>% 
  arrange(fips, year) %>% 
  ggplot(mapping = aes(x = med_income, y = calls_per_person)) +
  geom_point(aes(alpha = year), size = 2) +
  geom_path(aes(alpha = year), size = 1) +
  scale_alpha_continuous(breaks = seq(2010, 2017, 1), limits = c(2010, 2017))
```

```{r}
inner_join(nyc_2010, nyc_2017, by = "fips") %>%
  filter(total_pop.x > 2000, avg_rent.y > avg_rent.x) %>% 
  mutate(avg_rent.x = na_if(avg_rent.x, NaN)) %>% 
  drop_na(avg_rent.x, calls_per_person.x) %>% 
  ggplot(mapping = aes(x = avg_rent.x, y = calls_per_person.x)) +
  geom_point(aes(color = "2010")) +
  geom_point(aes(x = avg_rent.y, y = calls_per_person.y, color = "2017")) +
  geom_segment(aes(xend = avg_rent.y, yend = calls_per_person.y))

inner_join(nyc_2010, nyc_2017, by = "fips") %>%
  filter(total_pop.x > 2000, percent_white.y > percent_white.x) %>% 
  drop_na(percent_white.x, calls_per_person.x) %>% 
  ggplot(mapping = aes(x = percent_white.x, y = calls_per_person.x)) +
  geom_point(aes(color = "2010")) +
  geom_point(aes(x = percent_white.y, y = calls_per_person.y, color = "2017")) +
  geom_segment(aes(xend = percent_white.y, yend = calls_per_person.y))

inner_join(nyc_2010, nyc_2017, by = "fips") %>%
  filter(total_pop.x > 2000, med_income.y > med_income.x) %>% 
  drop_na(med_income.x, calls_per_person.x) %>% 
  ggplot(mapping = aes(x = med_income.x, y = calls_per_person.x)) +
  geom_point(aes(color = "2010")) +
  geom_point(aes(x = med_income.y, y = calls_per_person.y, color = "2017")) +
  geom_segment(aes(xend = med_income.y, yend = calls_per_person.y))

inner_join(nyc_2010, nyc_2017, by = "fips") %>%
  filter(total_pop.x > 2000, med_income.y > med_income.x) %>% 
  drop_na(med_income.x, calls_per_person.x) %>% 
  ggplot(mapping = aes(x = med_income.x, y = calls_per_person.x)) +
  geom_segment(
    aes(xend = med_income.y, yend = calls_per_person.y),
    arrow = arrow(length = unit(0.2, "cm"))
  )
```


```{r}
filter_data <- function(data){
  data %>% 
    filter(total_pop > 2000) %>% 
    drop_na(med_income)
}

filtered_2010 <- filter_data(nyc_2010)
filtered_2012 <- filter_data(nyc_2012)
filtered_2014 <- filter_data(nyc_2014)
filtered_2016 <- filter_data(nyc_2016)

filtered_2010 %>% 
  ggplot(aes(x = med_income, y = calls_per_person)) +
  geom_point(aes(color = "2010")) +
  geom_point(
    data = filtered_2012, 
    aes(x = med_income, y = calls_per_person, color = "2012")
  ) + 
  geom_point(
    data = filtered_2014, 
    aes(x = med_income, y = calls_per_person, color = "2014")
  ) + 
  geom_point(
    data = filtered_2016, 
    aes(x = med_income, y = calls_per_person, color = "2016")
  ) 
```

```{r}
filtered_2010 %>% 
  select(fips, med_income_2010 = med_income, calls_2010 = calls_per_person) %>% 
  inner_join(
    filtered_2012 %>% 
      select(fips, med_income_2012 = med_income, calls_2012 = calls_per_person),
    by = "fips"
  ) %>% 
  inner_join(
    filtered_2014 %>% 
      select(fips, med_income_2014 = med_income, calls_2014 = calls_per_person),
    by = "fips"
  ) %>% 
  inner_join(
    filtered_2016 %>% 
      select(fips, med_income_2016 = med_income, calls_2016 = calls_per_person),
    by = "fips"
  ) %>% 
  ggplot() +
  geom_segment(
    aes(
      x = med_income_2010, 
      y = calls_2010, 
      xend = med_income_2012, 
      yend = calls_2012
    ),
    color = "light gray"
  ) +
  geom_segment(
    aes(
      x = med_income_2012, 
      y = calls_2012, 
      xend = med_income_2014, 
      yend = calls_2014
    ),
    color = "light gray"
  ) +
  geom_segment(
    aes(
      x = med_income_2014, 
      y = calls_2014, 
      xend = med_income_2016, 
      yend = calls_2016
    ),
    color = "light gray"
  ) +
  geom_point(aes(x = med_income_2010, y = calls_2010, color = "2010")) + 
  geom_point(aes(x = med_income_2012, y = calls_2012, color = "2012")) + 
  geom_point(aes(x = med_income_2014, y = calls_2014, color = "2014")) + 
  geom_point(aes(x = med_income_2016, y = calls_2016, color = "2016")) +
  scale_color_manual(
    values = c("red3", "orange3", "yellow3", "springgreen4")
  ) +
  theme()
```

```{r}
inner_join(nyc_2010, nyc_2017, by = "fips") %>%
  filter(total_pop.x > 2000, med_income.y > med_income.x) %>% 
  drop_na(med_income.x, calls_per_person.x) %>% 
  ggplot(mapping = aes(x = med_income.x, y = calls_per_person.x)) +
  geom_segment(
    aes(xend = med_income.y, yend = calls_per_person.y),
    arrow = arrow(length = unit(0.2, "cm"))
  ) +
  geom_label(
    x = 200000, 
    y = 0.8, 
    label = "1 arrow = 1 census tract"
  ) +
  annotate(
    geom = "label",
    x = 203000,
    y = 0.5,
    label = "2010",
    color = "red",
    size = 3
  ) +
  annotate(
    geom = "label",
    x = 255000,
    y = 0.3,
    label = "2017",
    color = "red",
    size = 3
  ) +
  scale_x_continuous(labels = scales::dollar) +
  coord_fixed(ratio = 200000) +
  theme_classic() +
  labs(
    title = "311 Calls increase with Median Income from 2010 to 2017",
    subtitle = "Tracking 311 calls for Census Tracts in New York City",
    x = "Median Annual Income",
    y = "Calls per person",
    caption = "Source: NYC Open Data, American Fact Finder"
  )
```

# Looking at 2010 Alone

```{r}
single_year <- function(data){
  med_income <-
    data %>% 
    mutate(
      avg_med_income = weighted.mean(
        x = med_income, 
        w = total_pop, 
        na.rm = TRUE
      )
    ) %>% 
    pull(avg_med_income) %>% 
    unique()
  
  percent_white <-
    data %>% 
    mutate(
      avg_percent_white = weighted.mean(
        x = percent_white, 
        w = total_pop, 
        na.rm = TRUE
      )
    ) %>% 
    pull(avg_percent_white) %>% 
    unique()
  
  total_pop <-
    data %>% 
    mutate(
      tot_pop_nyc = sum(total_pop)
    ) %>% 
    pull(tot_pop_nyc) %>% 
    unique()
  
  total_calls <-
    data %>% 
    mutate(
      tot_calls = sum(num_calls)
    ) %>% 
    pull(tot_calls) %>% 
    unique()
  
  overview <- 
    tibble(
      `Total Population` = total_pop,
      `Median Income` = med_income,
      `Percent White` = percent_white,
      `Total # of 311 Calls` = total_calls
    ) 
}

single_year(nyc_2010)
single_year(nyc_2011)
```

## Plotting just about 2010

```{r}
nyc_2010 %>% 
  drop_na(med_income, calls_per_person) %>% 
  filter(calls_per_person < 1) %>% 
  ggplot(mapping = aes(x = med_income, y = calls_per_person)) +
  geom_point() +
  geom_smooth()

nyc_2010 %>% 
  drop_na(percent_white, calls_per_person) %>% 
  filter(calls_per_person < 1) %>% 
  ggplot(mapping = aes(x = percent_white, y = calls_per_person)) +
  geom_point() +
  geom_smooth()
```

## Plot just 2017

```{r}
nyc_2017 %>% 
  drop_na(med_income, calls_per_person) %>% 
  filter(calls_per_person < 1) %>% 
  ggplot(mapping = aes(x = med_income, y = calls_per_person)) +
  geom_point() +
  geom_smooth()

nyc_2017 %>% 
  drop_na(percent_white, calls_per_person) %>% 
  filter(calls_per_person < 1) %>% 
  ggplot(mapping = aes(x = percent_white, y = calls_per_person)) +
  geom_point() +
  geom_smooth()
```

```{r}
nyc_2010 %>% 
  ggplot(mapping = aes(x = med_income, y = calls_per_person)) +
  geom_smooth(aes(color = "2010"), se = FALSE, alpha = 0.2) +
  geom_smooth(data = nyc_2011, aes(color = "2011"), se = FALSE) +
  geom_smooth(data = nyc_2012, aes(color = "2012"), se = FALSE, alpha = 0.2) +
  geom_smooth(data = nyc_2013, aes(color = "2013"), se = FALSE, alpha = 0.2) +
  geom_smooth(data = nyc_2014, aes(color = "2014"), se = FALSE, alpha = 0.2) +
  geom_smooth(data = nyc_2015, aes(color = "2015"), se = FALSE, alpha = 0.2) +
  geom_smooth(data = nyc_2016, aes(color = "2016"), se = FALSE, alpha = 0.2) +
  geom_smooth(data = nyc_2017, aes(color = "2017"), se = FALSE, alpha = 0.2)
```


```{r}
nyc_2010 %>% 
  ggplot(mapping = aes(x = percent_white, y = calls_per_person)) +
  geom_smooth(aes(color = "2010"), se = FALSE) +
  geom_smooth(data = nyc_2011, aes(color = "2011"), se = FALSE) +
  geom_smooth(data = nyc_2012, aes(color = "2012"), se = FALSE) +
  geom_smooth(data = nyc_2013, aes(color = "2013"), se = FALSE) +
  geom_smooth(data = nyc_2014, aes(color = "2014"), se = FALSE) +
  geom_smooth(data = nyc_2015, aes(color = "2015"), se = FALSE) +
  geom_smooth(data = nyc_2016, aes(color = "2016"), se = FALSE) +
  geom_smooth(data = nyc_2017, aes(color = "2017"), se = FALSE)
```


## Plot change in median household income from 2010 to 2017

```{r}
tibble(
  fips = nyc_2010$fips,
  income_2010 = nyc_2010$med_income,
  calls_2010 = nyc_2010$calls_per_person,
  income_2017 = nyc_2017$med_income,
  calls_2017 = nyc_2017$calls_per_person,
  diff_income = income_2017 - income_2010,
  diff_calls = calls_2017 - calls_2010
) %>% 
  filter(calls_2010 < 1) %>% 
  drop_na(diff_income) %>% 
  ggplot(mapping = aes(x = diff_income, y = diff_calls)) +
  geom_point() +
  geom_smooth() +
  labs(
    title = "Difference in Calls per Person with Difference in Median Household Income"
  )
```

```{r}
tibble(
  fips = nyc_2010$fips,
  income_2010 = nyc_2010$med_income,
  calls_2010 = nyc_2010$calls_per_person,
  income_2017 = nyc_2017$med_income,
  calls_2017 = nyc_2017$calls_per_person,
  diff_income = income_2017 - income_2010,
  diff_calls = calls_2017 - calls_2010,
  percent_diff_income = diff_income / income_2010
) %>% 
  drop_na(diff_income) %>% 
  ggplot(mapping = aes(x = percent_diff_income, y = diff_calls)) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_point() +
  geom_smooth(color = "orangered2") +
  scale_x_continuous(
    breaks = seq(-0.25, 3, 0.25),
    labels = scales::percent
  ) +
  scale_y_continuous(breaks = seq(-0.8, 1, 0.2)) +
  coord_cartesian(ylim = c(-0.8, 1)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  labs(
    title = "Change in Calls per Person for Percent Change in Median Household Income",
    subtitle = "Change Measured between 2010 and 2017",
    x = "Percent Change in Median Household Income",
    y = "Change in 311 Calls per Person",
    caption = "Note: 1 Point = 1 Census Tract\nSource: US Census Bureau and NYC Open Data"
  )
```

## Plot change in percent white

```{r}
tibble(
  fips = nyc_2010$fips,
  white_2010 = nyc_2010$percent_white,
  calls_2010 = nyc_2010$calls_per_person,
  white_2017 = nyc_2017$percent_white,
  calls_2017 = nyc_2017$calls_per_person,
  diff_calls = calls_2017 - calls_2010,
  diff_white = white_2017 - white_2010
) %>% 
  drop_na(diff_white) %>% 
  ggplot(mapping = aes(x = diff_white, y = diff_calls)) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_point() +
  geom_smooth(color = "orangered2") +
  scale_x_continuous(
    breaks = seq(-0.4, 0.3, 0.1),
    labels = scales::percent
  ) +
  scale_y_continuous(breaks = seq(-1, 1.5, 0.2)) +
  coord_cartesian(xlim = c(-0.4, 0.3), ylim = c(-1, 1.5)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  labs(
    title = "Change in Calls per Person for Change in Percent White",
    subtitle = "Change Measured between 2010 and 2017",
    x = "Change in Percent White",
    y = "Change in 311 Calls per Person",
    caption = "Note: 1 Point = 1 Census Tract\nSource: US Census Bureau and NYC Open Data"
  )
```

## Just looking at noise complaints 

```{r}
police_calls_2010_file <- "~/Desktop/DCL/C01/police_calls/police_calls_2010.rds"
police_calls_2010 <- read_rds(police_calls_2010_file)
US_ALBERS <- "+proj=longlat +datum=WGS84 +no_defs"


just_heating <-
  police_calls_2010 %>% 
  select(Latitude, Longitude, `Complaint Type`, date) %>% 
  filter(!is.na(Latitude), !is.na(Longitude), `Complaint Type` == "HEATING") %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), crs = US_ALBERS) %>% 
  st_join(
    nyc_2010 %>% 
      select(-num_calls, -calls_per_person) %>% 
      st_as_sf() %>% 
      st_transform(crs = US_ALBERS)
  ) %>% 
  filter(!is.na(fips))

just_heating %>% 
  group_by(fips) %>% 
  mutate(
    heat_calls = n()
  ) %>% 
  ungroup() %>% 
  mutate(heat_calls_per_person = heat_calls / total_pop) %>% 
  ggplot(mapping = aes(x = percent_white, y = heat_calls_per_person)) +
  geom_point()
```

Getting correct colors for complaint type leaflet maps

```{r}
top_seven_complaints <-
  c(
    "Heating",
    "Noise",
    "General Construction",
    "Street Condition",
    "Plumbing",
    "Paint/Plaster",
    "Homeless Person Assistance",
    "Unsanitary Condition"
  )

top_calls <-
  police_calls_2010 %>% 
  mutate(
    complaints = str_to_lower(`Complaint Type`)
  ) %>%
  mutate(
    complaints = case_when(
      str_detect(complaints, "heat") ~ "heating",
      str_detect(complaints, "noise") ~ "noise",
      TRUE ~ complaints
    )
  ) %>% 
  select(Latitude, Longitude, complaints) %>% 
  filter(!is.na(Latitude), !is.na(Longitude)) %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), crs = US_ALBERS) %>% 
  st_join(
    nyc_2010 %>% 
      select(-num_calls, -calls_per_person) %>% 
      st_as_sf() %>% 
      st_transform(crs = US_ALBERS)
  ) %>% 
  filter(!is.na(fips))
```

Now I want to just look at heating complaints

```{r}
top_calls %>% 
  filter(complaints == "heating") %>% 
  group_by(fips) %>% 
  mutate(
    heat_calls = n()
  ) %>% 
  ungroup() %>% 
  mutate(heat_calls_per_person = heat_calls / total_pop) %>% 
  drop_na(percent_white, heat_calls_per_person) %>% 
  unique() %>% 
  ggplot(mapping = aes(x = percent_white, y = heat_calls_per_person)) +
  geom_point() +
  geom_smooth(color = "orangered2") +
  scale_x_continuous(
    breaks = seq(0, 1, 0.1),
    labels = scales::percent
  ) +
  scale_y_continuous(breaks = seq(-0.1, 0.5, 0.1)) +
  coord_cartesian(ylim = c(-0.1, 0.5)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  labs(
    title = "311 Heating Calls per Person for Percent White in 2010",
    x = "Percent White",
    y = "311 Heating Calls per Person",
    caption = "Note: 1 Point = 1 Census Tract\nSource: US Census Bureau and NYC Open Data"
  )
```

```{r}
top_calls %>% 
  filter(complaints == "heating") %>% 
  group_by(fips) %>% 
  mutate(
    heat_calls = n()
  ) %>% 
  ungroup() %>% 
  mutate(heat_calls_per_person = heat_calls / total_pop) %>% 
  drop_na(med_income, heat_calls_per_person) %>% 
  unique() %>% 
  ggplot(mapping = aes(x = med_income, y = heat_calls_per_person)) +
  geom_point() +
  geom_smooth(color = "orangered2") +
  scale_x_continuous(
    breaks = seq(0, 2.3e5, 25e3),
    labels = scales::dollar
  ) +
  scale_y_continuous(breaks = seq(0, 0.3, 0.05)) +
  coord_cartesian(ylim = c(0, 0.3)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  labs(
    title = "311 Heating Calls per Person for Median Income in 2010",
    x = "Median Income",
    y = "311 Heating Calls per Person",
    caption = "Note: 1 Point = 1 Census Tract\nSource: US Census Bureau and NYC Open Data"
  )
```

```{r}
police_calls_2017_file <- "~/Desktop/DCL/C01/police_calls/police_calls_2017.rds"
police_calls_2017 <- read_rds(police_calls_2017_file)

top_calls_2017 <-
  police_calls_2017 %>% 
  mutate(
    complaints = str_to_lower(`Complaint Type`)
  ) %>%
  mutate(
    complaints = case_when(
      str_detect(complaints, "heat") ~ "heating",
      str_detect(complaints, "noise") ~ "noise",
      TRUE ~ complaints
    )
  ) %>% 
  select(Latitude, Longitude, complaints) %>% 
  filter(!is.na(Latitude), !is.na(Longitude)) %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), crs = US_ALBERS) %>% 
  st_join(
    nyc_2017 %>% 
      select(-num_calls, -calls_per_person) %>% 
      st_as_sf() %>% 
      st_transform(crs = US_ALBERS)
  ) %>% 
  filter(!is.na(fips))
  
```

```{r}
top_calls_2017 %>% 
  filter(complaints == "heating") %>% 
  group_by(fips) %>% 
  mutate(
    heat_calls = n()
  ) %>% 
  ungroup() %>% 
  mutate(heat_calls_per_person = heat_calls / total_pop) %>% 
  drop_na(percent_white, heat_calls_per_person) %>% 
  unique() %>% 
  ggplot(mapping = aes(x = percent_white, y = heat_calls_per_person)) +
  geom_point() +
  geom_smooth(color = "orangered2") +
  scale_x_continuous(
    breaks = seq(0, 1, 0.1),
    labels = scales::percent
  ) +
  scale_y_continuous(breaks = seq(-0.1, 0.5, 0.1)) +
  coord_cartesian(ylim = c(-0.1, 0.5)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  labs(
    title = "311 Heating Calls per Person for Percent White in 2017",
    x = "Percent White",
    y = "311 Heating Calls per Person",
    caption = "Note: 1 Point = 1 Census Tract\nSource: US Census Bureau and NYC Open Data"
  )
```

```{r}
top_calls_2017 %>% 
  filter(complaints == "heating") %>% 
  group_by(fips) %>% 
  mutate(
    heat_calls = n()
  ) %>% 
  ungroup() %>% 
  mutate(heat_calls_per_person = heat_calls / total_pop) %>% 
  drop_na(med_income, heat_calls_per_person) %>% 
  unique() %>% 
  ggplot(mapping = aes(x = med_income, y = heat_calls_per_person)) +
  geom_point() +
  geom_smooth(color = "orangered2") +
  scale_x_continuous(
    breaks = seq(0, 2.3e5, 25e3),
    labels = scales::dollar
  ) +
  scale_y_continuous(breaks = seq(0, 0.2, 0.05)) +
  coord_cartesian(ylim = c(0, 0.2)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  labs(
    title = "311 Heating Calls per Person for Median Income in 2017",
    x = "Median Income",
    y = "311 Heating Calls per Person",
    caption = "Note: 1 Point = 1 Census Tract\nSource: US Census Bureau and NYC Open Data"
  )
```

# noise calls only

```{r}
top_calls %>% 
  filter(complaints == "noise") %>% 
  group_by(fips) %>% 
  mutate(
    noise_calls = n()
  ) %>% 
  ungroup() %>% 
  mutate(noise_calls_per_person = noise_calls / total_pop) %>% 
  drop_na(percent_white, noise_calls_per_person) %>% 
  unique() %>% 
  ggplot(mapping = aes(x = percent_white, y = noise_calls_per_person)) +
  geom_point() +
  geom_smooth(color = "orangered2") +
  scale_x_continuous(
    breaks = seq(0, 1, 0.1),
    labels = scales::percent
  ) +
  scale_y_continuous(breaks = seq(-0.5, 0.6, 0.1)) +
  coord_cartesian(ylim = c(-0.5, 0.6)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  labs(
    title = "311 Noise Calls per Person for Percent White in 2010",
    x = "Percent White",
    y = "311 Noise Calls per Person",
    caption = "Note: 1 Point = 1 Census Tract\nSource: US Census Bureau and NYC Open Data"
  )
```

```{r}
top_calls %>% 
  filter(complaints == "noise") %>% 
  group_by(fips) %>% 
  mutate(
    noise_calls = n()
  ) %>% 
  ungroup() %>% 
  mutate(noise_calls_per_person = noise_calls / total_pop) %>% 
  drop_na(med_income, noise_calls_per_person) %>% 
  unique() %>% 
  ggplot(mapping = aes(x = med_income, y = noise_calls_per_person)) +
  geom_point() +
  geom_smooth(color = "orangered2") +
  scale_x_continuous(
    breaks = seq(0, 2.3e5, 25e3),
    labels = scales::dollar
  ) +
  scale_y_continuous(breaks = seq(0, 0.35, 0.05)) +
  coord_cartesian(ylim = c(-0, 0.35)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  labs(
    title = "311 Noise Calls per Person for Median Income in 2010",
    x = "Median Income",
    y = "311 Noise Calls per Person",
    caption = "Note: 1 Point = 1 Census Tract\nSource: US Census Bureau and NYC Open Data"
  )
```

```{r}
top_calls_2017 %>% 
  filter(complaints == "noise") %>% 
  group_by(fips) %>% 
  mutate(
    noise_calls = n()
  ) %>% 
  ungroup() %>% 
  mutate(noise_calls_per_person = noise_calls / total_pop) %>% 
  drop_na(percent_white, noise_calls_per_person) %>% 
  unique() %>% 
  ggplot(mapping = aes(x = percent_white, y = noise_calls_per_person)) +
  geom_point() +
  geom_smooth(color = "orangered2") +
  scale_x_continuous(
    breaks = seq(0, 1, 0.1),
    labels = scales::percent
  ) +
  scale_y_continuous(breaks = seq(-0.6, 1.6, 0.4)) +
  coord_cartesian(ylim = c(-0.6, 1.6)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  labs(
    title = "311 Noise Calls per Person for Percent White in 2017",
    x = "Percent White",
    y = "311 Noise Calls per Person",
    caption = "Note: 1 Point = 1 Census Tract\nSource: US Census Bureau and NYC Open Data"
  )
```

```{r}
just_noise <- function(data){
  data %>% 
    filter(complaints == "noise") %>% 
    group_by(fips) %>% 
    mutate(
      noise_calls = n()
    ) %>% 
    ungroup() %>% 
    mutate(noise_calls_per_person = noise_calls / total_pop) %>% 
    select(fips, med_income, percent_white, noise_calls_per_person) %>% 
    drop_na(med_income, percent_white, noise_calls_per_person) %>% 
    arrange(fips) %>% 
    select(-geometry) %>% 
    unique()
}
```

```{r}
just_heating <- function(data){
  data %>% 
    filter(complaints == "heating") %>% 
    group_by(fips) %>% 
    mutate(
      heat_calls = n()
    ) %>% 
    ungroup() %>% 
    mutate(heat_calls_per_person = heat_calls / total_pop) %>% 
    select(fips, med_income, percent_white, heat_calls_per_person) %>% 
    drop_na(med_income, percent_white, heat_calls_per_person) %>% 
    arrange(fips) %>% 
    select(-geometry) %>% 
    unique()
}
```

```{r}
just_noise_2017 <- just_noise(top_calls_2017)

just_noise_2017 %>% 
  ggplot(mapping = aes(x = med_income, y = noise_calls_per_person)) +
  geom_point() +
  geom_smooth(color = "orangered2") +
  scale_x_continuous(
    breaks = seq(0, 2.3e5, 25e3),
    labels = scales::dollar
  ) +
  scale_y_continuous(breaks = seq(0, 0.35, 0.05)) +
  coord_cartesian(ylim = c(-0, 0.35)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  labs(
    title = "311 Noise Calls per Person for Median Income in 2017",
    x = "Median Income",
    y = "311 Noise Calls per Person",
    caption = "Note: 1 Point = 1 Census Tract\nSource: US Census Bureau and NYC Open Data"
  )
```

```{r}
just_noise_2010 <- just_noise(top_calls)
just_noise_2017 <- just_noise(top_calls_2017)

just_heating_2010 <- just_heating(top_calls)
just_heating_2017 <- just_heating(top_calls_2017)
```

```{r}
just_heating_2010 %>% 
  ggplot(mapping = aes(x = med_income, y = heat_calls_per_person)) +
  geom_point() +
  geom_smooth(color = "orangered2") +
  scale_x_continuous(
    breaks = seq(0, 2.3e5, 25e3),
    labels = scales::dollar
  ) +
  scale_y_continuous(breaks = seq(0, 0.3, 0.05)) +
  coord_cartesian(ylim = c(0, 0.3)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  labs(
    title = "311 Heating Calls per Person for Median Income in 2010",
    x = "Median Income",
    y = "311 Heating Calls per Person",
    caption = "Note: 1 Point = 1 Census Tract\nSource: US Census Bureau and NYC Open Data"
  )
```

```{r}
change_heating <- 
  just_heating_2010 %>% 
  inner_join(just_heating_2017, by = "fips") %>% 
  transmute(
    percent_diff_income = (med_income.y - med_income.x) / med_income.x,
    diff_white = percent_white.y - percent_white.x,
    diff_calls = heat_calls_per_person.y - heat_calls_per_person.x
  )

change_heating %>% 
  ggplot(mapping = aes(x = percent_diff_income, y = diff_calls)) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_point() +
  geom_smooth(color = "orangered2") +
  scale_x_continuous(
    breaks = seq(-0.25, 3, 0.25),
    labels = scales::percent
  ) +
  coord_cartesian(ylim = c(-0.2, 0.2)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  labs(
    title = "Change in Heating Calls per Person for\nPercent Change in Median Household Income",
    subtitle = "Change Measured between 2010 and 2017",
    x = "Percent Change in Median Household Income",
    y = "Change in 311 Calls per Person",
    caption = "Note: 1 Point = 1 Census Tract\nSource: US Census Bureau and NYC Open Data"
  )

change_heating %>% 
  ggplot(mapping = aes(x = diff_white, y = diff_calls)) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_point() +
  geom_smooth(color = "orangered2") +
  scale_x_continuous(
    breaks = seq(-0.4, 0.3, 0.1),
    labels = scales::percent
  ) +
  coord_cartesian(ylim = c(-0.2, 0.2)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  labs(
    title = "Change in Heating Calls per Person for Change in Percent White",
    subtitle = "Change Measured between 2010 and 2017",
    x = "Change in Percent White",
    y = "Change in 311 Heating Calls per Person",
    caption = "Note: 1 Point = 1 Census Tract\nSource: US Census Bureau and NYC Open Data"
  )
```

```{r}
change_noise <-
  just_noise_2010 %>% 
  inner_join(just_noise_2017, by = "fips") %>% 
  transmute(
    percent_diff_income = (med_income.y - med_income.x) / med_income.x,
    diff_white = percent_white.y - percent_white.x,
    diff_calls = noise_calls_per_person.y - noise_calls_per_person.x
  )

change_noise %>% 
  ggplot(mapping = aes(x = percent_diff_income, y = diff_calls)) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_point() +
  geom_smooth(color = "orangered2") +
  scale_x_continuous(
    breaks = seq(-0.25, 3, 0.25),
    labels = scales::percent
  ) +
  coord_cartesian(ylim = c(-0.05, 0.3)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  labs(
    title = "Change in Noise Calls per Person for\nPercent Change in Median Household Income",
    subtitle = "Change Measured between 2010 and 2017",
    x = "Percent Change in Median Household Income",
    y = "Change in 311 Noise Calls per Person",
    caption = "Note: 1 Point = 1 Census Tract\nSource: US Census Bureau and NYC Open Data"
  )

change_noise %>% 
  ggplot(mapping = aes(x = diff_white, y = diff_calls)) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_point() +
  geom_smooth(color = "orangered2") +
  scale_x_continuous(
    breaks = seq(-0.4, 0.3, 0.1),
    labels = scales::percent
  ) +
  coord_cartesian(ylim = c(-0.1, 0.3)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  labs(
    title = "Change in Noise Calls per Person for Change in Percent White",
    subtitle = "Change Measured between 2010 and 2017",
    x = "Change in Percent White",
    y = "Change in 311 Noise Calls per Person",
    caption = "Note: 1 Point = 1 Census Tract\nSource: US Census Bureau and NYC Open Data"
  )
```


## Mapping

```{r}
top_complaint_map <- function(call_data, nyc_data) {
  top_complaints <-
    call_data %>% 
    st_set_geometry(NULL) %>% 
    group_by(fips, complaints) %>% 
    summarize(num_call = sum(n())) %>% 
    group_by(fips) %>% 
    top_n(1, num_call) %>% 
    slice(1) %>% 
    ungroup() %>% 
    unique() %>% 
    arrange(fips) %>% 
    mutate(complaints = str_to_title(complaints)) %>% 
    mutate(
      complaints = if_else(
        complaints %in% top_seven_complaints,
        true = complaints,
        false = "Other"
      )
    ) %>% 
    left_join(nyc_data, by = "fips")
  
  labels_complaint <- 
    sprintf(
      "Top Complaint: %s",
      top_complaints$complaints
    ) %>% 
    lapply(htmltools::HTML)
  
  colors_complaints <- 
    colorFactor(
      palette = "Accent",
      domain = top_complaints$complaints 
      )
  
  leaflet() %>%  
    addProviderTiles(providers$CartoDB.Positron) %>% 
    setView(-73.95, 40.78, 11.4) %>% 
    addPolygons(
      data = top_complaints$geometry,
      weight = 1,
      fillOpacity = 1,
      color = colors_complaints(top_complaints$complaints),
      label = labels_complaint,
      highlightOptions = highlightOptions(
        color = "red", 
        weight = 3,
        bringToFront = TRUE
      )
    )
}
```

```{r}
top_complaint_map(top_calls, nyc_2010)
```

```{r}
police_calls_2011_file <- "~/Desktop/DCL/C01/police_calls/police_calls_2011.rds"
police_calls_2011 <- read_rds(police_calls_2011_file)

top_calls_2011 <-
  police_calls_2011 %>% 
  mutate(
    complaints = str_to_lower(`Complaint Type`)
  ) %>%
  mutate(
    complaints = case_when(
      str_detect(complaints, "heat") ~ "heating",
      str_detect(complaints, "noise") ~ "noise",
      TRUE ~ complaints
    )
  ) %>% 
  select(Latitude, Longitude, complaints) %>% 
  filter(!is.na(Latitude), !is.na(Longitude)) %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), crs = US_ALBERS) %>% 
  st_join(
    nyc_2011 %>% 
      select(-num_calls, -calls_per_person) %>% 
      st_as_sf() %>% 
      st_transform(crs = US_ALBERS)
  ) %>% 
  filter(!is.na(fips))
```

```{r}
top_complaint_map(top_calls_2011, nyc_2011)
```

