---
title: "Progress Report 2 Part 2"
author: "Natalie"
date: 2019-02-20
output: html_document
---

# Hello! Welcome to Natalie's scratchwork (because she organized her data really poorly before and made a mess out of the other RMarkdown files)

A little explanation for this file:

I am working with a large dataset that was taking forever to run, so I divided all my data up into smaller chunks to make it more manageable. This RMarkdown file is to have a workspace for these smaller data chunks, so I can save time in running and knitting!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(leaflet)

# Load in all data

 # Police call data
police_calls_2010_file <- "~/Desktop/DCL/C01/police_calls/police_calls_2010.rds"
police_calls_2011_file <- "~/Desktop/DCL/C01/police_calls/police_calls_2011.rds"
police_calls_2012_file <- "~/Desktop/DCL/C01/police_calls/police_calls_2012.rds"
police_calls_2013_file <- "~/Desktop/DCL/C01/police_calls/police_calls_2013.rds"
police_calls_2014_file <- "~/Desktop/DCL/C01/police_calls/police_calls_2014.rds"
police_calls_2015_file <- "~/Desktop/DCL/C01/police_calls/police_calls_2015.rds"
police_calls_2016_file <- "~/Desktop/DCL/C01/police_calls/police_calls_2016.rds"
police_calls_2017_file <- "~/Desktop/DCL/C01/police_calls/police_calls_2017.rds"

police_calls_2010 <- read_rds(police_calls_2010_file)
police_calls_2011 <- read_rds(police_calls_2011_file)
police_calls_2012 <- read_rds(police_calls_2012_file)
police_calls_2013 <- read_rds(police_calls_2013_file)
police_calls_2014 <- read_rds(police_calls_2014_file)
police_calls_2015 <- read_rds(police_calls_2015_file)
police_calls_2016 <- read_rds(police_calls_2016_file)
police_calls_2017 <- read_rds(police_calls_2017_file)

 # Rent data
rent_2010_file <- "~/Desktop/DCL/C01/rent/ACS_10_5YR_B25061_with_ann.csv"
rent_2011_file <- "~/Desktop/DCL/C01/rent/ACS_11_5YR_B25061_with_ann.csv"
rent_2012_file <- "~/Desktop/DCL/C01/rent/ACS_12_5YR_B25061_with_ann.csv"
rent_2013_file <- "~/Desktop/DCL/C01/rent/ACS_13_5YR_B25061_with_ann.csv"
rent_2014_file <- "~/Desktop/DCL/C01/rent/ACS_14_5YR_B25061_with_ann.csv"
rent_2015_file <- "~/Desktop/DCL/C01/rent/ACS_15_5YR_B25061_with_ann.csv"
rent_2016_file <- "~/Desktop/DCL/C01/rent/ACS_16_5YR_B25061_with_ann.csv"
rent_2017_file <- "~/Desktop/DCL/C01/rent/ACS_17_5YR_B25061_with_ann.csv"

 # Median household income data
income_2010_file <- 
  "~/Desktop/DCL/C01/median_household_income/ACS_10_5YR_B19013_with_ann.csv"
income_2011_file <- 
  "~/Desktop/DCL/C01/median_household_income/ACS_11_5YR_B19013_with_ann.csv"
income_2012_file <- 
  "~/Desktop/DCL/C01/median_household_income/ACS_12_5YR_B19013_with_ann.csv"
income_2013_file <- 
  "~/Desktop/DCL/C01/median_household_income/ACS_13_5YR_B19013_with_ann.csv"
income_2014_file <- 
  "~/Desktop/DCL/C01/median_household_income/ACS_14_5YR_B19013_with_ann.csv"
income_2015_file <- 
  "~/Desktop/DCL/C01/median_household_income/ACS_15_5YR_B19013_with_ann.csv"
income_2016_file <- 
  "~/Desktop/DCL/C01/median_household_income/ACS_16_5YR_B19013_with_ann.csv"
income_2017_file <- 
  "~/Desktop/DCL/C01/median_household_income/ACS_17_5YR_B19013_with_ann.csv"

 # Race data 
race_2010_file <- "~/Desktop/DCL/C01/race/ACS_10_5YR_B02001_with_ann.csv"
race_2011_file <- "~/Desktop/DCL/C01/race/ACS_11_5YR_B02001_with_ann.csv"
race_2012_file <- "~/Desktop/DCL/C01/race/ACS_12_5YR_B02001_with_ann.csv"
race_2013_file <- "~/Desktop/DCL/C01/race/ACS_13_5YR_B02001_with_ann.csv"
race_2014_file <- "~/Desktop/DCL/C01/race/ACS_14_5YR_B02001_with_ann.csv"
race_2015_file <- "~/Desktop/DCL/C01/race/ACS_15_5YR_B02001_with_ann.csv"
race_2016_file <- "~/Desktop/DCL/C01/race/ACS_16_5YR_B02001_with_ann.csv"
race_2017_file <- "~/Desktop/DCL/C01/race/ACS_17_5YR_B02001_with_ann.csv"

ny_census_tracts <-
  st_read(
    '~/Desktop/DCL/C01/gz_2010_36_140_00_500k/gz_2010_36_140_00_500k.shp'
  )
```

Wrangling data on rent:

```{r}
rent_floor_2010_2014 <-
  c(1000, 1250, 1500, 100, 150, 2000, 200, 250, 300, 350, 400, 450, 500,
    550, 600, 650, 700, 750, 800, 900, 0)

calculate_rent_prices_2010_2014 <- function(file_name) {
  file_name %>% 
  read_csv(
    col_names = TRUE,
    skip = 1,
    col_types = cols_only(
      Id = col_character(),
      `Estimate; Total: - Less than $100` = col_integer(),
      `Estimate; Total: - $100 to $149` = col_integer(),
      `Estimate; Total: - $150 to $199` = col_integer(),
      `Estimate; Total: - $200 to $249` = col_integer(),
      `Estimate; Total: - $250 to $299` = col_integer(),
      `Estimate; Total: - $300 to $349` = col_integer(),
      `Estimate; Total: - $350 to $399` = col_integer(),
      `Estimate; Total: - $400 to $449` = col_integer(),
      `Estimate; Total: - $450 to $499` = col_integer(),
      `Estimate; Total: - $500 to $549` = col_integer(),
      `Estimate; Total: - $550 to $599` = col_integer(),
      `Estimate; Total: - $600 to $649` = col_integer(),
      `Estimate; Total: - $650 to $699` = col_integer(),
      `Estimate; Total: - $700 to $749` = col_integer(),
      `Estimate; Total: - $750 to $799` = col_integer(),
      `Estimate; Total: - $800 to $899` = col_integer(),
      `Estimate; Total: - $900 to $999` = col_integer(),
      `Estimate; Total: - $1,000 to $1,249` = col_integer(),
      `Estimate; Total: - $1,250 to $1,499` = col_integer(),
      `Estimate; Total: - $1,500 to $1,999` = col_integer(),
      `Estimate; Total: - $2,000 or more` = col_integer()
    )
  ) %>% 
  gather(key = var_name, value = value, 2:ncol(.)) %>% 
  spread_(key = names(.)[1],value = 'value') %>% 
  mutate_at(
    vars(-var_name),
    ~ (. * rent_floor_2010_2014) / sum(.) %>% cumsum(.)
  ) %>% 
  gather(key = fips, value = rents, -var_name) %>% 
  select(-var_name) %>% 
  group_by(fips) %>% 
  summarize(
    avg_rent = sum(rents)
  ) 
}

rent_2010 <-
  calculate_rent_prices_2010_2014(rent_2010_file)

rent_2011 <-
  calculate_rent_prices_2010_2014(rent_2011_file)

rent_2012 <-
  calculate_rent_prices_2010_2014(rent_2012_file)

rent_2013 <-
  calculate_rent_prices_2010_2014(rent_2013_file)

rent_2014 <-
  calculate_rent_prices_2010_2014(rent_2014_file)

rent_floor_2015_2017 <-
  c(1000, 1250, 1500, 100, 150, 2000, 2500, 200, 250, 3000, 3500, 300,
    350, 400, 450, 500, 550, 600, 650, 700, 750, 800, 900, 0)

calculate_rent_prices_2015_2017 <- function(file_name) {
  file_name %>% 
  read_csv(
    col_names = TRUE,
    skip = 1,
    col_types = cols_only(
      Id = col_character(),
      `Estimate; Total: - Less than $100` = col_integer(),
      `Estimate; Total: - $100 to $149` = col_integer(),
      `Estimate; Total: - $150 to $199` = col_integer(),
      `Estimate; Total: - $200 to $249` = col_integer(),
      `Estimate; Total: - $250 to $299` = col_integer(),
      `Estimate; Total: - $300 to $349` = col_integer(),
      `Estimate; Total: - $350 to $399` = col_integer(),
      `Estimate; Total: - $400 to $449` = col_integer(),
      `Estimate; Total: - $450 to $499` = col_integer(),
      `Estimate; Total: - $500 to $549` = col_integer(),
      `Estimate; Total: - $550 to $599` = col_integer(),
      `Estimate; Total: - $600 to $649` = col_integer(),
      `Estimate; Total: - $650 to $699` = col_integer(),
      `Estimate; Total: - $700 to $749` = col_integer(),
      `Estimate; Total: - $750 to $799` = col_integer(),
      `Estimate; Total: - $800 to $899` = col_integer(),
      `Estimate; Total: - $900 to $999` = col_integer(),
      `Estimate; Total: - $1,000 to $1,249` = col_integer(),
      `Estimate; Total: - $1,250 to $1,499` = col_integer(),
      `Estimate; Total: - $1,500 to $1,999` = col_integer(),
      `Estimate; Total: - $2,000 to $2,499` = col_integer(),
      `Estimate; Total: - $2,500 to $2,999` = col_integer(),
      `Estimate; Total: - $3,000 to $3,499` = col_integer(),
      `Estimate; Total: - $3,500 or more` = col_integer()
    )
  ) %>% 
  gather(key = var_name, value = value, 2:ncol(.)) %>% 
  spread_(key = names(.)[1],value = 'value') %>% 
  mutate_at(
    vars(-var_name),
    ~ (. * rent_floor_2015_2017) / sum(.) %>% cumsum(.)
  ) %>% 
  gather(key = fips, value = rents, -var_name) %>% 
  select(-var_name) %>% 
  group_by(fips) %>% 
  summarize(
    avg_rent = sum(rents)
  ) 
}

rent_2015 <-
  calculate_rent_prices_2015_2017(rent_2015_file)

rent_2016 <-
  calculate_rent_prices_2015_2017(rent_2016_file)

rent_2017 <-
  calculate_rent_prices_2015_2017(rent_2017_file)
```

Wrangling data on median income:

```{r}
calculate_median_income <- function(filename) {
  filename %>% 
    read_csv(
      na = "-",
      skip = 2,
      col_names = FALSE,
      col_types = cols_only(
        X1 = col_character(),
        X4 = col_character()
      )
    ) %>% 
    transmute(
      fips = X1,
      med_income = if_else(
        condition = (X4 == "250,000+"),
        true = "250000",
        false = X4
      ) %>% 
        as.integer()
    )
}

income_2010 <- calculate_median_income(income_2010_file)
income_2011 <- calculate_median_income(income_2011_file)
income_2012 <- calculate_median_income(income_2012_file)
income_2013 <- calculate_median_income(income_2013_file)
income_2014 <- calculate_median_income(income_2014_file)
income_2015 <- calculate_median_income(income_2015_file)
income_2016 <- calculate_median_income(income_2016_file)
income_2017 <- calculate_median_income(income_2017_file)
```

Wrangling data on race:

```{r}
calculate_percent_white <- function(filename){
  filename %>% 
    read_csv(
      skip = 2,
      col_names = FALSE,
      col_types = cols_only(
        X1 = col_character(),
        X4 = col_double(),
        X6 = col_double()
      )
    ) %>% 
    transmute(
      fips = X1,
      percent_white = if_else(
        condition = (X4 == 0),
        true = NA_real_,
        false = X6 / X4
      )
    )
}

white_2010 <- calculate_percent_white(race_2010_file)
white_2011 <- calculate_percent_white(race_2011_file)
white_2012 <- calculate_percent_white(race_2012_file)
white_2013 <- calculate_percent_white(race_2013_file)
white_2014 <- calculate_percent_white(race_2014_file)
white_2015 <- calculate_percent_white(race_2015_file)
white_2016 <- calculate_percent_white(race_2016_file)
white_2017 <- calculate_percent_white(race_2017_file)
```

Now I'm going to join data to get one dataframe per year.

```{r}
nyc_2010 <-
  rent_2010 %>% 
  left_join(income_2010, by = "fips") %>% 
  left_join(white_2010, by = "fips")

nyc_2011 <-
  rent_2011 %>% 
  left_join(income_2011, by = "fips") %>% 
  left_join(white_2011, by = "fips")

nyc_2012 <-
  rent_2012 %>% 
  left_join(income_2012, by = "fips") %>% 
  left_join(white_2012, by = "fips")

nyc_2013 <-
  rent_2012 %>% 
  left_join(income_2013, by = "fips") %>% 
  left_join(white_2013, by = "fips")

nyc_2014 <-
  rent_2012 %>% 
  left_join(income_2014, by = "fips") %>% 
  left_join(white_2014, by = "fips")

nyc_2015 <-
  rent_2012 %>% 
  left_join(income_2015, by = "fips") %>% 
  left_join(white_2015, by = "fips")

nyc_2016 <-
  rent_2016 %>% 
  left_join(income_2016, by = "fips") %>% 
  left_join(white_2016, by = "fips")

nyc_2017 <-
  rent_2017 %>% 
  left_join(income_2017, by = "fips") %>% 
  left_join(white_2017, by = "fips")
```

Now need to join with tract data to get geometries!

```{r}
join_geometry <- function(df){
  df %>% 
    left_join(
      ny_census_tracts %>% 
        transmute(
          GEO_ID = as.character(GEO_ID)
        ), 
      by = c("fips" = "GEO_ID")
    )
}

nyc_2010 <- join_geometry(nyc_2010)
nyc_2011 <- join_geometry(nyc_2011)
nyc_2012 <- join_geometry(nyc_2012)
nyc_2013 <- join_geometry(nyc_2013)
nyc_2014 <- join_geometry(nyc_2014)
nyc_2015 <- join_geometry(nyc_2015)
nyc_2016 <- join_geometry(nyc_2016)
nyc_2017 <- join_geometry(nyc_2017)
```

Now for some interactive mapping!

```{r}
labels_2010 <- 
  sprintf(
  "Average Monthly Rent: $%g<br/>Median Household Income: $%g<br/>Percent White: %g",
  round(nyc_2010$avg_rent), 
  round(nyc_2010$med_income),
  round(100 * nyc_2010$percent_white)
  ) %>% lapply(htmltools::HTML)

white_vals <- 
  colorNumeric(
    palette = "Blues",
    domain = nyc_2010$percent_white
  )

leaflet() %>%  
  addProviderTiles(providers$Stamen.TonerLite) %>% 
  setView(-73.95, 40.78, 11) %>% 
  addPolygons(
    data = nyc_2010$geometry,
    weight = 1,
    fillOpacity = 1,
    color = white_vals(white_2010$percent_white),
    label = labels_2010,
    highlightOptions = highlightOptions(
      color = "red", 
      weight = 3,
      bringToFront = TRUE
    )
  ) %>%
  addLegend("bottomright", pal = white_vals, values = nyc_2010$percent_white,
    title = "Percentage White",
    labFormat = labelFormat(suffix = "%"),
    opacity = 1,
    na.label = NULL
  ) %>% 
  addCircleMarkers(
    lng = police_calls_2010$Longitude,
    lat = police_calls_2010$Latitude,
    radius = 0.1,
    label = police_calls_2010$`Complaint Type`
    )
```

